---
title: "Homework 5 - Dario Placencio"
output: html_document
date: "2023-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(digits = 5) 
```


### Consider data set PropertyFund.txt. We are interested in the prediction of variable Freq, i.e. the number of claims.

```{r}
# Load the data
property_fund <- read.table("PropertyFund.txt", header = TRUE)
```

### 1. Fit a Poisson regression using EntityType, AlarmCredit, and log of Coverage as predictors. Report your result.

```{r}
# Fit Poisson regression model with EntityType, AlarmCredit, and log of Coverage as predictors
poisson_model <- glm(Freq ~ EntityType + AlarmCredit + log(Coverage), data = property_fund, family = "quasipoisson")

# View the summary of the model
summary(poisson_model)
```
### 2. An analyst states that presumably the number of claims is proportional to the coverage amount, as a result, the variable Coverage should be treated as an exposure. Based on the results in 1, provide a counterargument for treating Coverage as an exposure.

Firstly, the coefficient for log(Coverage) in the model is positive and highly significant, suggesting that an increase in coverage is associated with an increase in the number of claims. This contradicts the assumption of proportionality between the number of claims and the coverage amount.

Secondly, the deviance residuals in the model suggest overdispersion, which means that the variance of the data is higher than what would be expected under a Poisson distribution. This indicates that there may be other factors beyond Coverage that are influencing the number of claims.

### 3. A main limitation of a Poisson regression is it does not accommodate overdispersion in the data. Provide an empirical evidence to support the presence of overdispersion.

One way to assess the presence of overdispersion is to compare the residual deviance to the residual degrees of freedom. If the residual deviance is much larger than the residual degrees of freedom, then there is evidence of overdispersion. In this case, the residual deviance is 6434.6 and the residual degrees of freedom is 5180, which suggests that there may be overdispersion present in the data. 

Additionally, the dispersion parameter estimated by the model is 1.7585, which is greater than 1, further suggesting that the data may be overdispersed.

### 4. Fit a negative binomial regression with the same predictors in the Poisson model. Report the results.

```{r}
library(MASS)

# Fit negative binomial model
model_nb <- glm.nb(Freq ~ EntityType + AlarmCredit + log(Coverage), data = property_fund)

# Print summary of results
summary(model_nb)
```
### 5. Poisson regression is a limiting case of negative binomial regression. Identify the parameter in the output in 3 to justify that the data support the negative binomial model over the Poisson model.

There is multiple approaches to compare the fitting level of the model. In question number 3, we used:

- Dispersion Parameter: For the Poisson distribution our value was well above 1 (with an actual value of 1.79) which suggest strong overdispersion for the model. For the negative binomial value is well different from 0 (with a value of 0.98) indicating the better fitting of this model.

- Deviance of the model: The deviance of the negative binomial model (3688) is significantly lower that the deviance of the poisson model (6434), close to 30% lower, suggesting than the first model is better fitting the data on this case.

### 6. Consider the Madison Metropolitan School District. The policyholder receives 10% alarm credit in premium, and has coverage amount of 100 million dollars.

```{r}
# Predict expected number of claims from Poisson model
poisson_preds <- predict(poisson_model, newdata = data.frame(EntityType = "School", AlarmCredit = "AC10%", Coverage = 100), type = "response")

# Predict expected number of claims from Negative Binomial model
nb_preds <- predict(model_nb, newdata = data.frame(EntityType = "School", AlarmCredit = "AC10%", Coverage = 100), type = "response")

```

#### 6.1 Compare the predicted expected number of claims from the Poisson and negative binomial models.

```{r}
# Print predicted expected number of claims from Poisson model
poisson_preds

# Print predicted expected number of claims from Negative Binomial model
nb_preds
```
The predicted expected number of claims from the Poisson model is 0.78371, while the predicted expected number of claims from the Negative Binomial model is 0.7926. These two predictions are very close to each other, suggesting that either model could be a reasonable choice for predicting the number of claims in the Madison Metropolitan School District. However, since the Negative Binomial model allows for greater flexibility in the variance, it may be a better choice if the variance of the number of claims is indeed larger than the mean.

#### 6.2 Compare the predicted probability that the policyholder will have more than 5 claims over the year from the Poisson and negative binomial models.

```{r}
# Calculate predicted probability for Poisson model
poisson_prob <- 1 - ppois(5, lambda=poisson_preds)
poisson_prob

# Calculate predicted probability for Negative Binomial model
nb_prob <- 1 - pnbinom(5, size=model_nb$theta, mu=nb_preds)
nb_prob
```
The predicted probability that the policyholder will have more than 5 claims over the year is much lower in the Poisson model (0.000165) compared to the negative binomial model (0.007729). This is expected since the negative binomial model allows for greater variability in the data and is more suitable when the mean and variance are not equal, which is the case for the given data. The Poisson model assumes that the mean and variance are equal, and hence it may underestimate the probability of extreme events.