## In Class Excercise - Dario Placencio

Set work environment
```{r}
#setwd("C:/Users/Dario/Documents/Github/ACTSCI657")
```

Read data txt
```{r}
data <- read.table("HealthData.txt", header = TRUE)
```

Take a subset of the data that contain observations with positive expenditure. 
```{r}
data <- data[data$erexp > 0, ]

```

Further split data into two parts, using observations in 2010 as training and observations in 2011 as test data.
```{r}
train <- data[data$year == 2010, ]
test <- data[data$year == 2011, ]
```

Import the package for gamma regression
```{r}
library(MASS)
```

Fit a gamma regression over the erexp variable. Use predictors chronic, race, and log of income in the model.
```{r}
glm(erexp ~ chronic + race + log(income), data = train, family = Gamma(link = "log")) -> model1
```

Report results
```{r}
summary(model1)
```

Identify a potential offset and/or weight variable. Refit the model by incorporating the offset/weight.
```{r}
glm(erexp ~ chronic + race + log(income), data = train, family = Gamma(link = "log"), weights = sqrt(ernum), offset = log(ernum)) -> model2
summary(model2)
```

Use mean as the predicted value. Predict the ER expenditure for individuals in the test data. 
Compare the prediction between above two models using mean squared prediction error (MSPE), i.e. ∑i=(yi−y^i)2. 
Indicated the preferred model suggested by MSPE

Use mean as the predicted value
```{r}
mean(test$erexp) -> mean_erexp
```

Predict the ER expenditure for individuals in the test data
```{r}
predict(model1, test, type='response') -> predicted_erexp1
```

Calculate the mean squared prediction error
```{r}
mean((mean_erexp - predicted_erexp1)^2) -> mspe1
mspe1
```

Use the fitted model to predict the ER expenditure for individuals in the test data
```{r}
predict(model2, test, type='response') -> predicted_erexp2
```

Calculate the mean squared prediction error
```{r}
mean((mean_erexp - predicted_erexp2)^2) -> mspe2
mspe2
```

Use the 80th percentile as the predicted value. Predict the ER expenditure for individuals in the test data. 
Compare the prediction between above two models using mean squared prediction error (MSPE), i.e. ∑i=(yi−y^i)2. 
Indicated the preferred model suggested by MSPE.

Use the 80th percentile as the predicted value
```{r}
quantile(train$erexp, 0.8) -> percentile_erexp
```

Predict the ER expenditure for individuals in the test data
```{r}
predict(model1, test, type='response') -> predicted_erexp3
```

Calculate the mean squared prediction error
```{r}
mean((percentile_erexp - predicted_erexp3)^2) -> mspe3
mspe3
```

Use the fitted model to predict the ER expenditure for individuals in the test data
```{r}
predict(model2, test, type='response') -> predicted_erexp4
```

Calculate the mean squared prediction error
```{r}
mean((percentile_erexp - predicted_erexp4)^2) -> mspe4
mspe4
```

Indicate the preferred model suggested by MSPE
```{r}
if (mspe1 < mspe2) {
  print("The preferred model is model1")
} else {
  print("The preferred model is model2")
}
```

Indicate the preferred model suggested by MSPE    
```{r}
if (mspe3 < mspe4) {
  print("The preferred model is model1")
} else {
  print("The preferred model is model2")
}
```